Incentivizing Exploration with Unbiased Histories

### 2019-02-01 COLT 2019 Submission/2level.tex
Lots of minor space-saving edits.
A bit more space-saving.
figure =>appendix
Some more space-saving.

%!TEX root = main.tex
\section{Warm-up: full-disclosure paths}
\label{sec:warmup}
We first consider a disclosure policy that reveals the full history in each round $t$, \ie $m_t=\SubH{t-1}$; we call it the \emph{full-disclosure policy}. The info-path for this policy is a simple path. We use this policy as a ``gadget" in our constructions. Hence, we formulate it slightly more generally:

@@

\begin{figure}[H]
\centering
\begin{tikzpicture}
 \filldraw[fill=blue!20!white]
 (0,2)--(10,2)--(10,3)--(0,3)--cycle;
  \filldraw[fill=red!20!white]
  (0,0)--(1,0)--(1,1)--(0,1)--cycle;
  \draw (0.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (1,0)--(2,0)--(2,1)--(1,1)--cycle;
  \draw (1.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (2,0)--(3,0)--(3,1)--(2,1)--cycle;
  \draw(2.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (3,0)--(4,0)--(4,1)--(3,1)--cycle;
  \draw(3.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (9,0)--(10,0)--(10,1)--(9,1)--cycle;
  \draw(9.5,1)--(5,2);
  \node at(5,0.5){$\cdots$};
  \node at(6,0.5){$\cdots$};
  \node at(7,0.5){$\cdots$};
  \node at(8,0.5){$\cdots$};
  \node at(5,2.5){all remaining rounds};
  \node at(0.5,0.5){$\fdpL$};
  \node at(1.5,0.5){$\fdpL$};
  \node at(2.5,0.5){$\fdpL$};
  \node at(3.5,0.5){$\fdpL$};
  \node at(9.5,0.5){$\fdpL$};
  \node at(-1,0.5){\textbf{Level 1}};
  \node at(-1,2.5){\textbf{Level 2}};
  \draw[->] (11,0)--(11,3);
  \node at(11.5,1.5)[ rotate=90]{Time};

@@

\caption{Info-graph for the 2-level policy. }
%Each red box in level 1 corresponds to a path connecting a set of
%  $T_G$ agents. The entire history in level 1 is then aggregated and
%  shown to each agent in level 2.}
\label{fig:2level}
\end{figure}

@@

\begin{theorem}\label{thm:2level}
The \2LEVEL with parameter $T_1 = T^{2/3}\,\log(T)^{1/3}$ achieves regret
\[ \reg(T) \leq O_K\left( T^{2/3}\, (\log T)^{1/3} \right).\]
\end{theorem}

@@

\begin{lemma}\label{lem:t1runs}
Suppose the info-graph contains $T_1$ full-disclosure paths of $\fdpL$ rounds each. Let $N_a$ be the number of samples of arm $a$ collected by all paths. Then with probability at
  least $1-\delta$, for all $a\in \A$,
  \[
    \left| N_a - \fdpN T_1\right| \leq \fdpL\cdot \sqrt{T_1 \log(2K/\delta) / 2}.
  \]
\end{lemma}

>>>>

%!TEX root = main.tex
\section{Warm-up: full-disclosure paths}
\label{sec:warmup}
We first consider a disclosure policy that reveals the full history in each round $t$, \ie $m_t=\SubH{t-1}$; we call it the \emph{full-disclosure policy}. The info-path for this policy is a simple path. We use this policy as a ``gadget" in our constructions. Hence, we formulate it more generally:

@@

%\iffalse
\begin{figure}[t]
\centering
\begin{tikzpicture}[scale =0.8]
 \filldraw[fill=blue!20!white]
 (0,2)--(10,2)--(10,3)--(0,3)--cycle;
  \filldraw[fill=red!20!white]
  (0,0)--(1,0)--(1,1)--(0,1)--cycle;
  \draw (0.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (1,0)--(2,0)--(2,1)--(1,1)--cycle;
  \draw (1.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (2,0)--(3,0)--(3,1)--(2,1)--cycle;
  \draw(2.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (3,0)--(4,0)--(4,1)--(3,1)--cycle;
  \draw(3.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (9,0)--(10,0)--(10,1)--(9,1)--cycle;
  \draw(9.5,1)--(5,2);
  \node at(5,0.5){$\cdots$};
  \node at(6,0.5){$\cdots$};
  \node at(7,0.5){$\cdots$};
  \node at(8,0.5){$\cdots$};
  \node at(5,2.5){all remaining rounds};
  \node at(0.5,0.5){$\fdpL$};
  \node at(1.5,0.5){$\fdpL$};
  \node at(2.5,0.5){$\fdpL$};
  \node at(3.5,0.5){$\fdpL$};
  \node at(9.5,0.5){$\fdpL$};
  \node at(-1,0.5){\textbf{Level 1}};
  \node at(-1,2.5){\textbf{Level 2}};
  \draw[->] (11,0)--(11,3);
  \node at(11.5,1.5)[ rotate=90]{Time};

@@

\caption{Info-graph for the 2-level policy. }
%Each red box in level 1 corresponds to a path connecting a set of
%  $T_G$ agents. The entire history in level 1 is then aggregated and
%  shown to each agent in level 2.}
\label{fig:2level}
\end{figure}
%\fi

@@

\begin{theorem}\label{thm:2level}
The \2LEVEL with parameter 
    $T_1 = \tilde{O}(T^{2/3})$ 
%    $T_1 = T^{2/3}\,\log(T)^{1/3}$ 
achieves regret
     $\tilde{O}_K (T^{2/3})$.
%\[ \reg(T) \leq O_K\left( T^{2/3}\, (\log T)^{1/3} \right).\]
\end{theorem}

@@

\begin{lemma}\label{lem:t1runs}
Suppose the info-graph contains $T_1$ full-disclosure paths of $\fdpL$ rounds each. Let $N_a$ be the number of samples of arm $a$ collected by all paths. Then with probability at
  least $1-\delta$,
  \[
    \left| N_a - \fdpN T_1\right| \leq \fdpL\cdot \sqrt{T_1 \log(2K/\delta) / 2},\quad
    \text{for all $a\in \A$}.
  \]
\end{lemma}

### 2019-02-01 COLT 2019 Submission/model.tex
Lots of minor space-saving edits.
Some more space-saving.

\xhdr{Unbiased subhistories.}
The \emph{subhistory} for a subset of rounds $S\subset [T]$ is defined as
\begin{align} \label{eq:model-subhistory}
    \SubH{S} = \left\{\; (s,a_s,r_s):\;s\in S \;\right\}.
\end{align}
$\SubH{[t-1]}$ is called the \emph{full history} at time $t$.
The \emph{outcome} for agent $t$ is the tuple $(t,a_t,r_t)$.

@@

\xhdr{Connection to multi-armed bandits.}
The special case when each message $m_t$ is an arm, and the $t$-th agent always chooses this arm, corresponds to a standard multi-armed bandit problem with IID rewards. Thus, regret in our problem can be directly compared to regret in the bandit problem with the same mean rewards $(\mu_a:\,a\in\A)$. Following the literature on bandits, we define the \emph{gap parameter} $\Delta$ as the difference between the largest and second largest mean rewards.%
\footnote{Formally, the second-largest mean reward is
    $\max_{a\in\A:\mu(a)<\mu^*} \mu(a)$,
where $\mu^* = \max_{a\in\A} \mu(a)$.
}
The gap parameter is not known to the principal (in our problem), or to the algorithm (in the bandit problem). Optimal regret rates for bandits with IID rewards are as follows \cite{bandits-ucb1,bandits-exp3,Lai-Robbins-85}:
\begin{align}\label{eq:model-OptRegret}
\reg(T) \leq O\left(\min\left(
    \sqrt{KT\log T},\; \tfrac{1}{\Delta} \log T
    \right)\right).
\end{align}
This regret bound can only be achieved using \emph{adaptive exploration}: \ie when exploration schedule is adapted to the observations. A simple example of \emph{non}-adaptive exploration is the \emph{explore-then-exploit} algorithm which samples arms uniformly at random for the first $N$ rounds, for some pre-set number $N$, then chooses one arm and sticks with it till the end. More generally,
\emph{exploration-separating} algorithms have a property that in each round $t$, either the choice of an arm does not depend on the observations so far, or the reward collected in this round is not used in the subsequent rounds. Any such algorithm suffers from $\Omega(T^{2/3})$ regret in the worst case.%
\footnote{The first explicit reference we know of is \cite{MechMAB-ec09,DevanurK09}, but this fact has been known in the community for much longer.}

@@

Throughout the paper, we use the standard concentration and anti-concentration inequalities: respectively, Chernoff Bounds and Berry-Esseen Theorem. The former states that that a sum of independent random variables converges to its expectation quickly. The latter states that the CDF of an appropriately scaled average of IID random variables converges to the CDF of the standard normal distribution pointwise. In particular, the average strays far enough from its expectation with some guaranteed probability. The theorem statements are moved to Appendix~\ref{sec:prelim}.

We use the notion of \emph{reward tape} to simplify the application of (anti-)concentration inequalities. This is a $K\times T$ random matrix with rows and columns corresponding to arms and rounds, respectively. For each arm $a$ and round $t$, the value in cell $(a,t)$ is drawn independently from Bernoulli distribution $\mD_a$. W.l.o.g., rewards in our model are defined by the rewards tape: namely, the reward for the $j$-th pull of arm $a$ is taken from the $(a,j)$-th entry of the reward matrix.

>>>>

\xhdr{Unbiased subhistories.}
The \emph{subhistory} for a subset of rounds $S\subset [T]$ is defined as
\begin{align*} %\label{eq:model-subhistory}
    \SubH{S} = \left\{\; (s,a_s,r_s):\;s\in S \;\right\}.
\end{align*}
$\SubH{[t-1]}$ is called the \emph{full history} at time $t$.
The \emph{outcome} for agent $t$ is the tuple $(t,a_t,r_t)$.

@@

\xhdr{Connection to multi-armed bandits.}
The special case when each message $m_t$ is an arm, and the $t$-th agent always chooses this arm, corresponds to a standard multi-armed bandit problem with IID rewards. Thus, regret in our problem can be directly compared to regret in the bandit problem with the same mean rewards $(\mu_a:\,a\in\A)$. Following the literature on bandits, we define the \emph{gap parameter} $\Delta$ as the difference between the largest and second largest mean rewards.%
\footnote{Formally, the second-largest mean reward is
    $\max_{a\in\A:\mu(a)<\mu^*} \mu(a)$,
where $\mu^* = \max_{a\in\A} \mu(a)$.
}
The gap parameter is not known to the principal (in our problem), or to the algorithm (in the bandit problem). Optimal regret rates for bandits with IID rewards are as follows \cite{bandits-ucb1,bandits-exp3,Lai-Robbins-85}:
\begin{align}\label{eq:model-OptRegret}
\reg(T) \leq O\left(\min\left(
    \sqrt{KT\log T},\; \tfrac{1}{\Delta} \log T
    \right)\right).
\end{align}
This regret bound can only be achieved using \emph{adaptive exploration}: \ie when exploration schedule is adapted to the observations. A simple example of \emph{non}-adaptive exploration is the \emph{explore-then-exploit} algorithm which samples arms uniformly at random for the first $N$ rounds, for some pre-set $N$, then chooses one arm and sticks with it till the end. Any such algorithm suffers from $\Omega(T^{2/3})$ regret in the worst case.%
\footnote{More generally, this lower bound applies to \emph{exploration-separating} algorithms, where in each round $t$, either the choice of an arm does not depend on the observations so far, or the reward collected in this round is not used in the subsequent rounds. The first explicit reference we know of is \cite{MechMAB-ec09,DevanurK09}, but this fact has been known in the community for much longer.}

@@

Throughout the paper, we use the standard concentration and anti-concentration inequalities: resp., Chernoff Bounds and Berry-Esseen Theorem (see Appendix~\ref{sec:prelim}). The former states that that a sum of independent random variables converges to its expectation quickly. The latter states that the CDF of an appropriately scaled average of IID random variables converges to the CDF of the standard normal distribution pointwise. In particular, the average strays far enough from its expectation with some guaranteed probability.
%The theorem statements are moved to Appendix~\ref{sec:prelim}.

We use the notion of \emph{reward tape} to simplify applications of (anti-)concentration inequalities. This is a $K\times T$ random matrix with rows and columns corresponding to arms and rounds. For each arm $a$ and round $t$, the value in cell $(a,t)$ is drawn independently from Bernoulli distribution $\mD_a$. W.l.o.g., rewards in our model are defined by the rewards tape: the reward for the $j$-th pull of arm $a$ is taken from cell $(a,j)$ of the reward tape.
